---
title: "EDS230; Function 1"
author: "Bailey Jørgensen"
format: html
editor: visual
---

Write a function that computes energy produced from a photovoltaic system if you know the average annual solar radiation. I will use the following formula:

    $E = A * r * H * PR$

    * E is energy (kWh),
    * A is the solar panel area (m2)
    * r is panel yield (0-1) (manufacture efficiency - usually around 0.2),
    * PR is performance ratio (0-1) (accounting for site factors that impact efficiency usually around 0.75) 
    * H is annual average solar radiation (kWh)

```{r}
# Function to calculate energy from a photovoltaic system
calculate_energy <- function(A, r, H, PR) {
  # Calculate energy (E)
  E <- A * r * H * PR
  return(E)
}
```

Now that I've created the function, I will use some random numbers/fake data, to test this function out!

```{r}
# Random number for: solar panel area (m2)
A <- 10

# Random number for: panel yield (efficiency)
r <- 0.2

# Random number for: annual average solar radiation (kWh)
H <- 1500

# Random number for: performance ratio
PR <- 0.75

# Calculate energy in this fake scenario!
energy <- calculate_energy(A, r, H, PR)
print(paste("Energy produced:", energy, "kWh"))
```

